Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-02 overleaf/related.tex
Changes to results and related work.

HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY/MOTION circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All~\ana{add citation!}, however, it is not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- specifically, it is well-known that vectorization and polyhedral compilation 
does not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.

>>>>

HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All~\ana{add citation!}, however, it is not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- concretely, it is well-known that vectorization and polyhedral compilation 
do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.

### 2022-05-02 overleaf/results.tex
Changes to results and related work.

\subsection{Discussion}
\ana{Ana todo: Discuss why we are slower than HyCC --- because of ad-hoc div-and-conquer optimizations. Have to summarize what Ben and I figured out yesterday.}

\ana{If we can add that our cost model works great (it does!), that would be great but probably won't have time...}

>>>>

\subsection{Comparison with MOTION-native Inner Product and Discussion}

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i] + mult\_unsimdified[i];}. 
We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!

Therefore, MOTION's compiler performs analysis that informs circuit generation. In the above example, 
MOTION does the standard divide-and-conquer accumulation. The example also illustrates the limitation of higher-level AST analysis --- while 
the analysis was able to detect the associative accumulation when operator {\sf += } was used, it was unable to do so when an equivalent 
code was used. We conjecture that MPC source, as a more straight-forward representation will
not only enable detection of general associative loops, but also allow for program synthesis to increase opportunities for divide-and-conquer
parallelization~\cite{Farzan:2021}; we leave this as future work.

The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
Again, we estimate we are about an order of magnitude slower. This is likely due to the same issue as Inner Product --- the computation of min can be optimized.

%\ana{Ana todo: Discuss why we are slower than HyCC --- because of ad-hoc div-and-conquer optimizations. Have to summarize what Ben and I figured out yesterday.}

%\ana{If we can add that our cost model works great (it does!), that would be great but probably won't have time...}
